---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# shinytune

<!-- badges: start -->
<!-- badges: end -->

The goal of shinytune is to make it easy to explore `tune` objects, similar to
`shinystan`.

To do this I need to:

* [ ] Figure out exactly what `tune` is producing
* [ ] Think about some sensible summaries/visualisations/explorations that could
be applied to that
* [ ] (Optionally) Compare to `shinystan` for reference (I don't want to anchor
too strongly to it, though)

## Exploring `tune`

Firstly, I need to figure out what  `tune` object actually contains.

Let's create one following the Getting Started guide on
the [tune
website](https://tidymodels.github.io/tune/articles/getting_started.html).

```{r}
library(tidymodels)
library(tune)
```

### Tune objects

Let's just create the whole set of outputs for the Getting Started vignette.

Firstly we'll set up the data:

```{r}
library(AmesHousing)

ames <- make_ames()

set.seed(4595)
data_split <- initial_split(ames, strata = "Sale_Price")
ames_train <- training(data_split)
ames_test  <- testing(data_split)
```

Then the baseline recipe

```{r}
ames_rec <- recipe(Sale_Price ~ Longitude + Latitude, data = ames_train) %>% 
  step_log(Sale_Price, base = 10) %>% 
  step_ns(Longitude, deg_free = tune("long df")) %>% 
  step_ns(Latitude,  deg_free = tune("lat df"))

ames_rec
```

Then update the parameters to use a better function with a wider range: 

```{r}
ames_param <- ames_rec %>% 
  parameters() %>% 
  update(
    `long df` = spline_degree(), 
    `lat df` = spline_degree()
  )

ames_param
```

Then we'll set up the (grid) search space of values for these parameters:

```{r}
spline_grid <- grid_max_entropy(ames_param, size = 10)
spline_grid
```

Then our (linear) model:

```{r}
lm_mod <- linear_reg() %>% 
    set_engine("lm")

lm_mod
```

Then we'll set up the cross validation scheme to search over:

```{r}
set.seed(2453)
cv_splits <- vfold_cv(ames_train, v = 10, strata = "Sale_Price")
```

Then finally we'll do the tuning using `tune_grid()`:

```{r cache = TRUE}
ames_res <- tune_grid(
    ames_rec,
    model = lm_mod,
    resamples = cv_splits,
    grid = spline_grid
)
ames_res
```

